---
title:  "Networks: Project 2"
author: "Dmitry Donetskov"
# date:   {today}

output: 
  pdf_document:
    toc: true
    number_sections: true
    toc_depth: 2
    fig_width: 7
    fig_height: 7
    fig_caption: true
        
bibliography: project_2.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(igraph)
library(orca)

library(RColorBrewer)
```

# Problem Statement

Select an undirected labeled network with at least 50 nodes and average degree
at least 4. Determine the graphlet spectra for its nodes. On their basis
construct a dissimilarity between nodes and use it to cluster the nodes.

# Network

The data is the network of American football games between Division IA colleges 
during regular season Fall 2000, as compiled by M. Girvan and M. Newman 
(@Girvan7821).

```{r data_prep, echo=FALSE}
net <- read_graph("data/football.gml", format = "gml")
```

## Basic Network Characteristics 

The nodes represent web pages, each page describes some Wikipedia social norm. 
The links are the HTTP reference from one page to another.

It's a directed graph by its nature, no weights assigned to the links.

|Property               | Value |
|:----------------------|:-----------------|
| Vertices              | `r gorder(net)`                                      |
| Edges                 | `r as.character(gsize(net))`                         |
| Directed              | `r ifelse(is.directed(net), "Yes", "No")`            |
| Weighted              | `r ifelse(is.weighted(net), "Yes", "No")`            |
| Average degree        | `r round(2*gsize(net)/gorder(net), 1)`               |
| Diameter              | `r diameter(net, directed = TRUE, weights = NULL)`   |
| Acyclic               | `r ifelse(is.dag(net), "Yes", "No")`                 |
| Edge density          | `r as.character(edge_density(net))`                  |
| Average Path Length   | `r as.character(mean_distance(net, directed = TRUE))`|
| Transitivity (global) | `r as.character(transitivity(net))`                  |


```{r net_viz, echo=FALSE}
plot(net, vertex.color="green", layout=layout_with_kk, vertex.label = NA, vertex.size = 5)
```
# Analysis

## Getting Orbits

```{r data_prep2, echo=TRUE}
# convert to the 'edgelist' format required by orca
net_edgelist <- as.data.frame(as_edgelist(net))
net_edgelist <- transform(net_edgelist, V1 = as.integer(net_edgelist$V1), V2 = as.integer(net_edgelist$V2))
```


```{r calc, echo=TRUE}
# convert to the 'edgelist' format required by orca
net_orbits        <- count4(net_edgelist)
net_orbits_scaled <- scale(net_orbits)     # scale the data

print('Found orbits for the nodes are')
print(net_orbits)
# 
```

## H-Clustering

```{r hclust, echo=TRUE, fig.height = 7, fig.width = 7, fig.align = "center"}
heatmap(net_orbits_scaled, cexRow = 0.3) 

# Ward Hierarchical Clustering
distances  <- dist(net_orbits_scaled, method = "euclidean") # distance matrix
hclust_fit <- hclust(distances, method="ward.D") 
plot(hclust_fit, cex = 0.3) # display dendogram
k = 10
# draw dendogram with red borders around the k clusters 
rect.hclust(hclust_fit, k = k, border="red")

clusters2 <- cutree(hclust_fit, k = k) # cut tree into k clusters
```

## k-Means Clustering

```{r kmeans, echo=TRUE, fig.height = 7, fig.width = 7, fig.align = "center"}
k = 10 

kmeans_fit <- kmeans(net_orbits_scaled, k)

# Cluster Plot against 1st 2 principal components
library(cluster) 
clusplot(net_orbits_scaled, kmeans_fit$cluster, color = TRUE, shade = TRUE, labels = 2, lines=0)

# Centroid Plot against 1st 2 discriminant functions
library(fpc)
plotcluster(net_orbits_scaled, kmeans_fit$cluster)
```

## Visualization (Clustering Coloured)

```{r net_viz_hclust, echo=TRUE, fig.height = 10, fig.width = 10, fig.align = "center"}
# color the nodes
pal = colorRampPalette(brewer.pal(11,"Spectral"))(length(unique(clusters2)))
node_colors <- pal[clusters2]
V(net)$color <- node_colors

coords_fr = layout.fruchterman.reingold(net, weights=E(net)$weight)
# igraph plot options
igraph.options(vertex.size=8, edge.width=0.75) 
# plot network
plot(net, layout=coords_fr, vertex.color=V(net)$color)
```
```{r net_viz_kmeans, echo=TRUE, fig.height = 10, fig.width = 10, fig.align = "center"}
# color the nodes
pal = colorRampPalette(brewer.pal(11,"Spectral"))(length(unique(kmeans_fit$cluster)))
node_colors <- pal[kmeans_fit$cluster]
V(net)$color <- node_colors

# coords_fr = layout.fruchterman.reingold(net, weights=E(net)$weight)
# igraph plot options
igraph.options(vertex.size=8, edge.width=0.75) 
# plot network
plot(net, layout=coords_fr, vertex.color=V(net)$color)
```

# References

