---
title: 'Networks: Project 2'
author: "Dmitry Donetskov"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
    toc: yes
    toc_depth: 2
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
bibliography: project_2.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(igraph)
library(orca)

library(RColorBrewer)
```

# Problem Statement

Select an undirected labeled network with at least 50 nodes and average degree
at least 4. Determine the graphlet spectra for its nodes. On their basis
construct a dissimilarity between nodes and use it to cluster the nodes.

# Network

The data is the network of American football games between Division IA colleges 
during regular season Fall 2000, as compiled by M. Girvan and M. Newman 
(@Girvan7821). To quote

>The network we look at is a representation of the schedule of Division I games
for the 2000 season: vertices in the graph represent teams (identified by their
college names) and edges represent regular-season games between the two teams
they connect. What makes this network interesting is that it incorporates a
known community structure. The teams are divided into conferences containing
around 8–12 teams each. Games are more frequent between members of the same
conference than between members of different conferences, with teams playing an
average of about seven intraconference games and four interconference games in
the 2000 season.

Let's check if the clustering based on the graphlets can reveal the true
structure of the network.


```{r data_prep, echo=FALSE}
net <- read_graph("data/football.gml", format = "gml")
```

## Basic Network Characteristics 

|Property               | Value |
|:----------------------|:-----------------|
| Vertices              | `r gorder(net)`                                      |
| Edges                 | `r as.character(gsize(net))`                         |
| Directed              | `r ifelse(is.directed(net), "Yes", "No")`            |
| Weighted              | `r ifelse(is.weighted(net), "Yes", "No")`            |
| Average degree        | `r round(2*gsize(net)/gorder(net), 1)`               |
| Diameter              | `r diameter(net, directed = TRUE, weights = NULL)`   |
| Acyclic               | `r ifelse(is.dag(net), "Yes", "No")`                 |
| Edge density          | `r as.character(edge_density(net))`                  |
| Average Path Length   | `r as.character(mean_distance(net, directed = TRUE))`|
| Transitivity (global) | `r as.character(transitivity(net))`                  |


```{r net_viz, echo = TRUE, dev='svg', fig.height = 10, fig.width = 10, fig.align = "center"}
coords_fr = layout.fruchterman.reingold(net, weights=E(net)$weight)

# igraph plot options
igraph.options(vertex.size=8, edge.width=0.75) 
plot(net, vertex.color="green", layout=coords_fr, vertex.size = 5)
```
# Analysis

The [orca](https://cran.r-project.org/web/packages/orca/orca.pdf) package is used to find associations of nodes with the 4-node orbits.

## Getting Orbits

```{r data_prep2, echo=TRUE}
# convert to the 'edgelist' format required by orca
net_edgelist <- as.data.frame(as_edgelist(net))
net_edgelist <- transform(net_edgelist, V1 = as.integer(net_edgelist$V1), V2 = as.integer(net_edgelist$V2))
```

Printing the graphlet spectra.

```{r calc, echo=TRUE}
# convert to the 'edgelist' format required by orca
net_orbits        <- count4(net_edgelist)
net_orbits_scaled <- scale(net_orbits)     # scale the data

print('Found orbits for the nodes are')
print(net_orbits)
# 
```

## H-Clustering

I use the euclidian distance for hierarchical clustering and made the cut to
form ten partitions.

```{r hclust, echo=TRUE, dev='svg', fig.height = 7, fig.width = 7, fig.align = "center"}
heatmap(net_orbits_scaled, cexRow = 0.3) 

# Ward Hierarchical Clustering
distances  <- dist(net_orbits_scaled, method = "euclidean") # distance matrix
hclust_fit <- hclust(distances, method="ward.D") 
plot(hclust_fit, cex = 0.3) # display dendogram
k = 10
# draw dendogram with red borders around the k clusters 
rect.hclust(hclust_fit, k = k, border="red")

clusters2 <- cutree(hclust_fit, k = k) # cut tree into k clusters
```

## k-Means Clustering

The same number of centers (10) was used for k-Means clustering to check later
if results of two different clustering methods will match.

```{r kmeans, echo=TRUE, dev='svg', fig.height = 7, fig.width = 7, fig.align = "center"}
k = 10 

kmeans_fit <- kmeans(net_orbits_scaled, k)

# Cluster Plot against 1st 2 principal components
library(cluster) 
clusplot(net_orbits_scaled, kmeans_fit$cluster, color = TRUE, shade = TRUE, labels = 2, lines=0)

# Centroid Plot against 1st 2 discriminant functions
library(fpc)
plotcluster(net_orbits_scaled, kmeans_fit$cluster)
```

## Visualization (Hierarchical  Clustering)

The network coloured based on the clusters defined during hierarchical clustering.

```{r net_viz_hclust, echo=TRUE, dev='svg', fig.height = 10, fig.width = 10, fig.align = "center"}
# color the nodes
pal = colorRampPalette(brewer.pal(11,"Spectral"))(length(unique(clusters2)))
node_colors <- pal[clusters2]
V(net)$color <- node_colors

#coords_fr = layout.fruchterman.reingold(net, weights=E(net)$weight)
# igraph plot options
igraph.options(vertex.size=8, edge.width=0.75) 
# plot network
plot(net, layout=coords_fr, vertex.color=V(net)$color)
```
## Visualization (k-Means)

The network coloured based on the clusters defined during k-Means clustering.

```{r net_viz_kmeans, echo=TRUE, dev='svg', fig.height = 10, fig.width = 10, fig.align = "center"}
# color the nodes
pal = colorRampPalette(brewer.pal(11,"Spectral"))(length(unique(kmeans_fit$cluster)))
node_colors <- pal[kmeans_fit$cluster]
V(net)$color <- node_colors

# coords_fr = layout.fruchterman.reingold(net, weights=E(net)$weight)
# igraph plot options
igraph.options(vertex.size=8, edge.width=0.75) 
# plot network
plot(net, layout=coords_fr, vertex.color=V(net)$color)
```

# Findings

TBC

# References

