---
title: 'Networks: Project 3'
author: "Dmitry Donetskov"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
    toc: yes
    toc_depth: 2
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
bibliography: project_3.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# chunk option dev="svg" produces very large vector graphics files
knitr::opts_chunk$set(dev="pdf")

# chunk option dev="png" is the default raster graphics format for HTML output
#knitr::opts_chunk$set(dev="png")

library(igraph)
library(RColorBrewer)
```

# Problem Statement

Make a CUG test on a network of your choice.

# Network

The data is the network of American football games between Division IA colleges 
during regular season Fall 2000, as compiled by M. Girvan and M. Newman 
(@Girvan7821). To quote

>The network we look at is a representation of the schedule of Division I games
for the 2000 season: vertices in the graph represent teams (identified by their
college names) and edges represent regular-season games between the two teams
they connect. What makes this network interesting is that it incorporates a
known community structure. The teams are divided into conferences containing
around 8–12 teams each. Games are more frequent between members of the same
conference than between members of different conferences, with teams playing an
average of about seven intraconference games and four interconference games in
the 2000 season.


By description of the network we can assume there are components in it which 
have more connections within and between them.

Let's check with CUG 

```{r data_prep, echo=FALSE}
net <- read_graph("data/football.gml", format = "gml")
```

## Basic Network Characteristics 

|Property               | Value |
|:----------------------|:-----------------|
| Vertices              | `r gorder(net)`                                      |
| Edges                 | `r as.character(gsize(net))`                         |
| Directed              | `r ifelse(is.directed(net), "Yes", "No")`            |
| Weighted              | `r ifelse(is.weighted(net), "Yes", "No")`            |
| Average degree        | `r round(2*gsize(net)/gorder(net), 1)`               |
| Diameter              | `r diameter(net, directed = TRUE, weights = NULL)`   |
| Acyclic               | `r ifelse(is.dag(net), "Yes", "No")`                 |
| Edge density          | `r as.character(edge_density(net))`                  |
| Average Path Length   | `r as.character(mean_distance(net, directed = TRUE))`|
| Transitivity (global) | `r as.character(transitivity(net))`                  |

The network coloured according to which division each club belongs to.

```{r net_viz, echo=TRUE, fig.height = 10, fig.width = 10, fig.align = "center"}
pal = colorRampPalette(brewer.pal(11,"Spectral"))(length(unique(V(net)$value)))
node_colors  <- pal[V(net)$value + 1]
V(net)$color <- node_colors

coords_fr = layout.fruchterman.reingold(net, weights=E(net)$weight)

# igraph plot options
igraph.options(vertex.size=8, edge.width=0.75) 
plot(net, layout=coords_fr, vertex.size = 5, vertex.color=V(net)$color)
```

# Analysis

```{r cug, echo=TRUE, warning=FALSE, message=FALSE}
library(statnet)

# converting from the 'igraph' format to the 'network' one
net2 <- network(as.matrix(as_adjacency_matrix(net, type="both", names=TRUE, sparse=FALSE)), 
                matrix.type="adjacency", directed=FALSE)

rSize  <- cug.test(net2, centralization, FUN.arg=list(FUN=betweenness), mode="graph", cmode="size")
rEdges <- cug.test(net2, centralization, FUN.arg=list(FUN=betweenness), mode="graph", cmode="edges")
rDyad  <- cug.test(net2, centralization, FUN.arg=list(FUN=betweenness), mode="graph", cmode="dyad.census")

#names(rSize)
#rSize

# Aggregate results
Betweenness <- c(rSize$obs.stat, rEdges$obs.stat, rDyad$obs.stat)
PctGreater  <- c(rSize$pgteobs,  rEdges$pgteobs,  rDyad$pgteobs)
PctLess     <- c(rSize$plteobs,  rEdges$plteobs,  rDyad$plteobs)
report      <- cbind(Betweenness, PctGreater, PctLess)
rownames(report) <- c("Size","Edges","Dyads")
```

```{r cug2, echo=TRUE}
report

par(mfrow=c(1,3))
plot(rSize,  main="Betweenness \nConditioned on Size" )
plot(rEdges, main="Betweenness \nConditioned on Edges" )
plot(rDyad,  main="Betweenness \nConditioned on Dyads" )
par(mfrow=c(1,1))
```

# Findings

The CUG analysis shows that the network has unusually high level of
centralization for the given number of nodes and that relfects the nature of
network (that connections between teams within divisions are more likely than
between teams from different divisions).
But this level of centralization is not uncommon (and rather a bit low) for the
given number of edges and the given distribution of dyads.

# References

